---
title: "Multilevel contingency table model on simulated fake data"
author: "Martin Johnsson"
date: "2017-08-05"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

We will demonstrate the model by simulating fake data and fitting the model to it.


## Fake data simulation

The package contains a simple function that lets us simulate data from the assumed model
and a related model that assumes a fixed difference in probability between
the two groups.

```{r}
data <- sim_tables(alphaA = 0.5, betaA = 5,
                   alphaB = 5, betaB = 0.5,
                   delta = NULL)
```

The data frame look like this, which is the format that the model fitting expects,
with two additional columns containing the true probabilities for each group.

```{r}
data
```

The simulation function looks like this -- it first samples the probabilities from beta
distributions, and then samples the data given those probabilites.

```{r}
sim_tables
```



## Model fitting

The rstan model has been compiled on package installation. We can run it with this function:

```{r, collapse = TRUE}
model <- fit_model(data)
```

Sampling uses the stan default of four chains of 2000 samples each, 1000 of which are discarded
as warmup, and no thinning. The only difference is that, similar to rstanarm, the adapt_delta parameter
has been increased to 0.95. This makes stan take smaller steps when exploring the posterior,
sacrificing some speed for lower risk of running into numerical problems.

You can change the number of chains and samples by passing ```chains```, ```iter```, and ```warmup```
to the function as you would in rstan.

As you can see, the ```fit_model``` function is really a very thin wrapper around rstan:

```{r}
fit_model
```



## Results

The result is a stanfit object, and we can then use all the nice rstan summary and plotting
functions on it. Here, for example, are the estimated probabilities for both groups and each
of the ten tables, their credible intervals, and the diagnostic potential scale reduction factor:

```{r}
print(model, pars = c("pA", "pB"))
```

Here are the hyperparameters:

```{r}
print(model, pars = c("alphaA", "alphaB", "betaA", "betaB"))
```

And the estimated differences in probability between the groups for each table:

```{r}
print(model, pars = c("delta", "average_delta"))
```

We can compare this to the simulated values:

```{r}
data$pB_true - data$pA_true
(true_average <- mean(data$pB_true - data$pA_true))
```

(Note: the ```rstan::``` and ```ggplot2::``` are there to avoid using the ```library``` function in 
the vignette. In a normal script, you can just load rstan and ggplot2.)

```{r}
rstan::plot(model, pars = c("delta", "average_delta")) + 
  ggplot2::geom_vline(xintercept = true_average)
```

